---
title: "Chimp Personality (30 Dec)"
author: "Noah Stetson"
date: "`r Sys.Date()`"
format: html
editor: visual
---

### Data analysis outline for chimp personality (with beta version of survey)

### Goals:

1.  Load library and data

    -   This is connected to GitHub

2.  Clean data

3.  Survey data

    -   Inter-item correlation: Do certain test items correlate with each other? Do the test items correlate as expected?

    -   Inter-total correlation: Do certain personality scores correlate with each other?

    -   Inter-rater correlation: See how the 2 raters (AS and RR) correlate with how they scores the chimps

**Later:**

1.  Factors impacting how people rate personality

    1.  Time knowing the chimp

    2.  Focal subject vs non-focal

    3.  US researcher vs Ugandan research staff

    4.  Research project

2.  Compare "dominance" score on survey to other measures of dominance

    1.  Aggression

    2.  Pant grunts

    3.  Elo ratings

    4.  David's score

------------------------------------------------------------------------

### Load library and data

#### Library

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(psych)
library(readr)
library(corrplot)
library(irr)
library(lares)
```

#### Data

```{r}
Dec28 <- read_csv("https://github.com/nwstetson/ChimpPersonality/blob/main/Data/28Dec2022_Raw.csv", show_col_types = FALSE)  # survey data taken from Qualtrics on December 28th 2022
```

------------------------------------------------------------------------
